A memory frontier for complex synapses

### 2013-05-14 NIPS/SynapseNIPS.tex
wrote initial SNR section

\section{Mathematical setup}\label{sec:setup}

@@

This results in the following SNR
%
\begin{equation}\label{eq:SNRcalc}
  \snr(t) = \sqrt{N}\prn{2 f\pot f\dep} \eq \prn{\M\pot-\M\dep} \e^{rt\frg} \w.
\end{equation}
%
We will frequently refer to this function as the memory curve.

@@

>>>>

\section{Synaptic models and their memory curves}\label{sec:setup}

In this section, we will describe the class of models of synaptic plasticity that we are studying and how we quantify their performance in memory storage.
In the subsequent sections, we will find upper bounds on this performance.

@@

This results in the following SNR
%
\begin{equation}\label{eq:SNRcalc}
  \snr(t) = \sqrt{N}\prn{2 f\pot f\dep} \eq \prn{\M\pot-\M\dep} \e^{rt\frg} \w,
\end{equation}
%
where the weight of the synapse when it is in its $i$'th state is given by the corresponding element of the column vector $\w$.
We will frequently refer to this function as the memory curve.

@@

In this section, we will find some upper bounds on certain properties of the memory curve discussed in the previous section, namely its initial value and the area under it.
We will use these bounds to determine an upper bound on the SNR at finite times in the next section.

\subsection{Initial SNR}\label{sec:initial}

Now we will discuss the SNR at $t=0$:
%
\begin{equation}\label{eq:init}
  \snr(0) = \sqrt{N}\prn{2 f\pot f\dep} \eq \prn{\M\pot-\M\dep} \w.
\end{equation}
%
We will find an upper bound on this quantity for \emph{all} possible models and also find the model that saturates this bound.

A useful quantity is the equilibrium probability flux between two disjoint sets of states, $\CA$ and $\CB$:
%
\begin{equation}\label{eq:flux}
  \F_{\CA\CB} = \sum_{i\in\CA} \sum_{j\in\CB} r \eq_i \frg_{ij}.
\end{equation}
%
The initial SNR is closely related to the flux between the states with $\w_i=-1$ and those with $\w_j=+1$:
%
\begin{equation}\label{eq:initflux}
  \snr(0) \leq \frac{4\sqrt{N}\F_{-+}}{r}\,.
\end{equation}
%
This inequality becomes an equality if potentiation never decreases the synaptic weight and depression never increases it, which should be a property of any sensible model.

To maximise this flux, potentiation from a weak state must be guaranteed to end in a strong state, and depression must do the reverse.
AN example of such a model is shown in \autoref{fig:maxinit}(\ref{fig:max_init_pot},\ref{fig:max_init_dep}).
These models have a property known as ``lumpability'' (see \cite[\S6.3]{kemeny1960finite} for the discrete time version and \cite{burke1958markovian,Ball1993Lumpability} for continuous time).
They are completely equivalent (\ie have the same memory curve) as a two state model with transition probabilities equal to 1, as shown in \autoref{fig:maxinit}(\ref{fig:binary_det}).

This two state model has the equilibrium distribution $\eq=(f\dep,f\pot)$ and its flux is given by $\F_{-+} = rf\pot f\dep$. 
This is maximized when $f\pot=f\dep=\half$, leading to the upper bound:
%
\begin{equation}\label{eq:maxinit}
  \snr(0) \leq \sqrt{N}.
\end{equation}
%

\begin{figure}
 \begin{center}
 \begin{myenuma}
  \item\hp\aligntop{\includegraphics[width=0.25\linewidth]{max_init_pot.svg}}\label{fig:max_init_pot}\hp
  \item\aligntop{\includegraphics[width=0.25\linewidth]{max_init_dep.svg}}\label{fig:max_init_dep}\hp
  \item\aligntop{\includegraphics[width=0.1\linewidth]{binary_det.svg}}\label{fig:binary_det}
  \end{myenuma}
 \end{center}
  \caption{Synaptic models that maximize initial SNR. 
  (\ref{fig:max_init_pot}) For potentiation, all transitions starting from a weak state lead to a strong state, and the probabilities for all transitions leaving a given weak state sum to 1.
  (\ref{fig:max_init_pot}) Depression is similar to potentiation, but with strong and weak interchanged. 
  (\ref{fig:binary_det}) The equivalent two state model, with transition probabilities under potentiation and depression equal to one.
  }\label{fig:maxinit}
\end{figure}

\subsection{Area}\label{sec:area}

Now consider the area under the memory curve:
%
\begin{equation}\label{eq:area}
  A = \int_0^\infty\!\!\dr t \, \snr(t).
\end{equation}
%
We will find an upper bound on this quantity as well as the model that saturates it.
